---
title: "Basic {tidymodels} example"
author: "Adam BUshman"
output: html_document
---

# Basic {tidymodels} examples

In my learning path of {tidymodels}, I'm creating this resource as a start guide for creating many of the core model types. They will be small, getting started examples that can be repurposed into grander projects.


## Setup

```{r}
#| echo: FALSE

library('tidyverse')
library('tidymodels')
library('broom')
```


## Linear Regression

Using the car prices data set from {modeldata}, wherein we'll look to model the price of a car models based on discrete and continuous variables like mileage, cylinders, body type, etc.

```{r}
data_lmr <- modeldata::car_prices
```

```{r}
lmr_mod <-
  linear_reg() %>%
  set_mode("regression") %>%
  set_engine("lm")

lmr_mod
```

We initially call the model we want to use. Then, we set the mode (only one choice for this model). Then, we set the engine (only one choice for this model).

```{r}
fit_lmr <-
  lmr_mod %>%
  fit(Price ~ ., data = data_lmr)

fit_lmr
```

With the model defined, we now "fit" (define which variables are independent and which are dependent). 

```{r}
fit_lmr %>%
  pluck("fit") %>%
  summary()
```

With the model fit to the data, we can inspect the summary easily enough by extracting it from the fit object.

```{r}
# Fit summary put in a tibble
fit_lmr %>% 
  tidy()

# Summary statistics put in a tibble
fit_lmr %>%
  glance()

# Appends the fitted and residual values to the original data
fit_lmr %>%
  pluck("fit") %>%
  augment()
```

We should now tidy up the output for easier use in plotting and table generation.

