---
title: "Zoo Animal Classification"
description: "Leveraging decision trees in {tidymodels} to classify zoo animals"
author: "Adam Bushman"
format: html
---


# Introduction

Data sourced from Kaggle,
[Zoo Animal Classification](https://www.kaggle.com/datasets/uciml/zoo-animal-classification/data), accessed January 19th, 2024. Downloaded extracts are saved within this repository and loaded for purpose of modeling.


# Analysis prep

```{r}

# Loading libraries
library('tidyverse')
library('tidymodels')

```

```{r}

# Defining directory reference
here::i_am("projects/decision-tree-classification/zoo-animal-classes/model.qmd")

# Load data into session
zoo <- read.csv("zoo.csv")
class <- read.csv("class.csv")

```


# Basic data exploration

```{r}

# Understand fiends, 
glimpse(zoo)

summary(zoo)

```

# Alter dataset

```{r}

# Turn `class_type` to a factor
zoo <- 
  zoo %>%
  select(-animal_name) %>%
  mutate(
    across(hair:fins, ~ as.logical(.)), 
    across(tail:catsize, ~ as.logical(.)), 
    class_type = factor(class_type)
  )

```


# Model prep

```{r}

# Defining a split
split <- rsample::initial_split(zoo)

# Training and testing sets
z_train <- rsample::training(split)
z_test <- rsample::testing(split)

```

# Defining a model, recipe, and workflow

```{r}

# Model
z_mod <- decision_tree(tree_depth = tune()) %>%
  set_mode("classification") %>%
  set_engine("rpart")

# Recipe
z_rec <- recipe(class_type ~ ., data = z_train)

# Workflow
z_wfl <- workflow() %>%
  add_model(z_mod) %>%
  add_recipe(z_rec)

```

# Setting up the tuning details

```{r}

# Tuning grid
z_grid <- expand.grid(tree_depth = seq(1,10))

# Setting up the tuning
z_tune <- tune_grid(
  z_wfl, 
  resamples = rsample::bootstraps(z_train, times = 5), 
  grid = z_grid, 
  metrics = metric_set(accuracy)
)

```


# Exploring the results

```{r}

# Extracting results from the tuned grid
results <- 
  z_tune %>%
  unnest(.metrics)

# Plot accuracy metric
ggplot(results, aes(x = tree_depth, y = .estimate)) +
  geom_point() +
  stat_smooth() +
  scale_x_continuous(
    breaks = results$tree_depth
  )

```

# Constructing a model with max depth of 4

```{r}

# Official model
z_mod_o <- 
  update(
    z_mod, 
    tree_depth = 4
  )

# Official workflow
z_wfl_o <- 
  workflow() %>% 
  add_model(z_mod_o) %>% 
  add_recipe(z_rec)

# Fitting the training data
z_fit <-
  z_wfl_o %>%
  fit(data = z_train)

```

# Evaluating the model

```{r}

# Predicting the testing data based on training
z_pred <- 
  z_fit %>%
  predict(new_data = z_test)

# Assembling results 
results <- tibble(
  predicted = z_pred$.pred_class, 
  observed = z_test$class_type
)

# Comparing results
results %>%
  yardstick::metrics(
    truth = observed, 
    estimate = predicted
  )

```
